---
title: "p8105_hw6_ng3005"
author: "Nelson Gaillard"
date: "2025-12-01"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(broom)

set.seed(1)
```

## Problem 2

```{r, cache = TRUE}
data("weather_df")

boot_res =
  replicate(5000, {
    boot_dat = weather_df |> 
      sample_frac(replace = TRUE)
    
    fit = lm(tmax ~ tmin + prcp, data = boot_dat)
    
    r2 = glance(fit) |> 
      pull(r.squared)
    
    coefs = tidy(fit)
    
    b1 = coefs |> 
      filter(term == "tmin") |> 
      pull(estimate)
    
    b2 = coefs |> 
      filter(term == "prcp") |> 
      pull(estimate)
    
    ratio = b1 / b2
    
    c(r2 = r2, b1_over_b2 = ratio)
  })

boot_df = as_tibble(t(boot_res))

boot_df |> 
  pivot_longer(cols = everything(),
               names_to = "quantity",
               values_to = "value") |> 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ quantity, scales = "free_x") +
  labs(x = "Bootstrap estimate", y = "Count")
```

The distribution of the ratio of beta_1 to beta_2 appears to be left-skewed and unimodal with a majority of the data concentrated closer to the right side of the plot. 

The distribution of r2 appears to be relatively symmetric and unimodal with a median bootstrap estimate hovering between 0.940 and 0.945. 

```{r}
boot_df |> 
  summarize(
    r2_low = quantile(r2, 0.025), 
    r2_high = quantile(r2, 0.975),
    ratio_low = quantile(b1_over_b2, 0.025),
    ratio_high = quantile(b1_over_b2, 0.975)
  ) |> 
  knitr::kable(digits = 2)
```

Based on the table, the 95% CI for the ratio of beta_1 to beta_2 is (-281.31, -124.80). The 95% CI for r2 is (0.93, 0.95). 

## Problem 3

```{r}
birth_df =
  read_csv("data/birthweight.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  mutate(
    babysex = as.factor(babysex),
    mrace = as.factor(mrace),
    frace = as.factor(frace),
    malform = as.factor(malform))

main_model = lm(
  bwt ~ gaweeks + blength + bhead + ppbmi + wtgain + smoken + parity + babysex + mrace + fincome,
  data = birth_df)

birth_df |> 
  modelr::add_predictions(main_model) |> 
  modelr::add_residuals(main_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(
    x = "Fitted birthweight (g)", 
    y = "Residuals", 
    title = "Residuals vs Fitted Values for Birthweight Model"
  )
```

This model contains the outcome of interest, `bwt`, and the two exposures of interest, `gaweeks` and `blength` in addition to various relevant covariates--ones that clinically affect the outcome of interest. It is important to adjust for these convariates in the hypothesized model to avoid spurious associations due to confounding.

#### Comparison to additional models

```{r}
cv_splits = modelr::crossv_mc(birth_df, n = 50)

main_effects =
  lm(bwt ~ blength + gaweeks, data = birth_df)

interaction = 
  lm(bwt ~ bhead * blength * babysex, data = birth_df)

cv_splits |> 
  mutate(
    fit_main = map(train, ~ lm(main_model, data = as_tibble(.x))),
    fit_main_effects = map(train, ~ lm(main_effects, data = as_tibble(.x))),
    fit_int = map(train, ~ lm(interaction, data = as_tibble(.x))),
    
    rmse_main = map2_dbl(fit_main, test, ~ {
      test_df = as_tibble(.y)
      preds = predict(.x, newdata = test_df)
      truth = test_df |> pull(bwt)
      sqrt(mean((preds - truth)^2))
    }),
    
    rmse_main_effects = map2_dbl(fit_main_effects, test, ~ {
      test_df = as_tibble(.y)
      preds = predict (.x, newdata = test_df)
      truth = test_df |> pull(bwt)
      sqrt(mean((preds - truth)^2))
    }),
    
    rmse_interaction = map2_dbl(fit_int, test, ~ {
      test_df = as_tibble(.y)
      preds = predict(.x, newdata = test_df)
      truth = test_df |> pull(bwt)
      sqrt(mean((preds - truth)^2))
    })
  ) |> 
  
  summarize(
    mean(rmse_main),
    mean(rmse_main_effects),
    mean(rmse_interaction)
  ) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "mean_rmse"
  ) |> 
  knitr::kable()
```

This table shows the `mean_rmse` for each of the three models we're comparing--`main_model`, `main_effects`, and `interaction`. Based on the analysis, `main_model` has the lowest mean rmse, meaning this model's predictions are closest to the true birthweight compared to the other models. On the other hand, `main_effects`, the model containing only the variables `gaweek` and `blength`, has the greatest rmse. This indicates that this model does not fit the data as well as the other two. For `interaction`, mean rmse decreased notably, but not to the same extent as `main_model` which adjusts for numerous covariates.
